---
title: "Joining Air Quality and Zip Code"
author: "Gustavo Garcia-Franceschini"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries}
library(sf)
library(sfheaders)
library(tidyverse)
```


# Doing some shape math

What's going on here is I take the center of each zip code area, and figure out which UHF neighborhood the that point in inside of. This will then tell me which zip codes belong to which UHF neighborhoods, and will allow us to join the air quality and disease datasets.

One shortcoming of this method is that we don't actually know if the entire zip code area is inside the UHF neighborhood; just the center.

```{r shapes}
air_shapes = read_sf(dsn = "raw_shapes/", layer = "UHF42") %>%
  filter(id != "0")

zip_codes = read_csv("raw_data/nyc-zip-codes.csv") %>%
  mutate(zip = as.character(ZipCode)) %>%
  select(-ZipCode) %>%
  janitor::clean_names()

zip_shapes = read_sf(dsn = "raw_shapes/", layer = 'tl_2019_us_zcta510') %>% 
  rename(zip = ZCTA5CE10) %>%
  inner_join(zip_codes) %>%
  st_transform(crs= st_crs(air_shapes)) %>%
  st_point_on_surface()
```

```{r contains}
#source: https://dewey.dunnington.ca/post/2022/profiling-point-in-polygon-joins-in-r/
uhf_contains_zips = st_join(zip_shapes, air_shapes, join = st_within) %>%
  sf_to_df(fill = T)  %>%
  select(zip, id)
```

# Loading data

```{r loading_and_writing}
df_air = read_csv("raw_data/air_quality.csv") %>%
  janitor::clean_names() %>%
  select(-message) %>%
  mutate(id = as.character(geo_join_id)) %>%
  filter(geo_type_name == "UHF42")

joined_datasets = df_air %>%
  full_join(uhf_contains_zips) %>%
  full_join(zip_codes)

write_csv(joined_datasets, "cleaned_data/air_quality_zips.csv")
```

# Joining data with lat,log

```{r}
df_air_geo = 
  read_csv("cleaned_shapes/zip_shapes.csv") |> 
  select(zip, intptlat10, intptlon10) |> 
  left_join(read_csv("cleaned_data/air_quality_zips.csv"), by = "zip") |> 
  mutate(lat = intptlat10,
         lon = intptlon10) |> 
  select(-intptlat10, -intptlon10)
```

