---
title: "Project Shape Cleaning"
author: "Gustavo Garcia-Franceschini"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries}
library(sf)
library(sfheaders)
library(tidyverse)
```

# Air Quality Data

```{r air_data}
df_air = read_csv("raw_data/air_quality.csv") %>%
  janitor::clean_names() %>%
  select(-message) %>%
  mutate(id = as.character(geo_join_id)) %>%
  filter(geo_type_name == "UHF42")
```

# Shape Files

```{r air_shapes}
air_shapes = read_sf(dsn = "raw_shapes/", layer = "UHF42")
```

```{r zip_shapes_and_projections}
zip_codes = read_csv("raw_data/nyc-zip-codes.csv") %>%
  mutate(zip = as.character(ZipCode))

zip_shapes = read_sf(dsn = "raw_shapes/", layer = 'tl_2019_us_zcta510') %>% 
  rename(zip = ZCTA5CE10) %>%
  inner_join(zip_codes) %>%
  st_transform(crs= st_crs(air_shapes)) %>%
  mutate(x_center = map_dbl(geometry, ~st_point_on_surface(.x)[[1]]),
         y_center = map_dbl(geometry, ~st_point_on_surface(.x)[[2]])) %>%
  sf_to_df(fill = T)  %>%
  janitor::clean_names()

air_shapes = air_shapes %>%
  sf_to_df(fill = T) %>%
  janitor::clean_names() %>%
  mutate(geo_join_id = as.numeric(id)) %>%
  filter(id != "0")
```

# Example map

The map shows the 42 UHF neighborhoods (colors), with the zip code bounds drawn over the colors.

```{r example_map}
ggplot() + geom_polygon(data = air_shapes, 
                        aes(x = x, y = y, group = id), 
                        fill = "white", color = "black") + 
  theme_void() + geom_point(data = zip_shapes, 
                            aes(x = x_center, y = y_center, 
                                color = neighborhood)) + 
  theme(legend.position = "none")
```

# Joining Air Data with UHF Shape files 

Before joining, I think you should make `df_air` smaller (by choosing a specific year and name to track).

```{r}
#df_air_shapes = df_air %>%
#  inner_join(air_shapes, by = "id")
```

# Joining with zip code data

Joining these will be hard because the neighborhoods don't have the same names. We don't have to join them for the maps, but we'd need to join them for regression.

```{r why_joining_will_be_hard}
sort(unique(df_air$geo_place_name))

sort(unique(zip_codes$Neighborhood))
```

# Exporting air and zip shape files

```{r shape_export}
write_csv(air_shapes, "cleaned_shapes/air_shapes.csv")

write_csv(zip_shapes, "cleaned_shapes/zip_shapes.csv")
```

