---
title: "Project Shape Cleaning"
author: "Gustavo Garcia-Franceschini"
date: "`r Sys.Date()`"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries}
library(sf)
library(sfheaders)
library(tidyverse)
```

# Air Quality Data

```{r air_data}
#source: https://data.cityofnewyork.us/Environment/Air-Quality/c3uy-2p5r
df_air = read_csv("raw_data/air_quality.csv") %>%
  janitor::clean_names() %>%
  select(-message) %>%
  mutate(id = as.character(geo_join_id)) %>%
  filter(geo_type_name == "UHF42")
```

# Shape Files

This code involves: 

- loading the shape files using the `sf` package

- loading the `zip_codes` dataset so we can know which of the zip codes in the `zip_shapes` dataset are from NYC (as opposed to other parts of New York).

- transforming the `zip_shapes` dataset from regular coordinates to EPSG:2263

- changing everything to regular datasets for easier plotting

```{r air_shapes}
#source: https://github.com/nycehs/NYC_geography/tree/master
air_shapes = read_sf(dsn = "raw_shapes/", layer = "UHF42")
```

```{r zip_shapes_and_projections}
#source: https://github.com/erikgregorywebb/nyc-housing/blob/master/Data/nyc-zip-codes.csv
zip_codes = read_csv("raw_data/nyc-zip-codes.csv") %>%
  mutate(zip = as.character(ZipCode)) %>%
  select(-ZipCode) %>%
  janitor::clean_names()

#source: https://catalog.data.gov/dataset/tiger-line-shapefile-2019-2010-nation-u-s-2010-census-5-digit-zip-code-tabulation-area-zcta5-na 
zip_shapes = read_sf(dsn = "raw_shapes/", layer = 'tl_2019_us_zcta510') %>% 
  rename(zip = ZCTA5CE10) %>%
  inner_join(zip_codes) %>%
  st_transform(crs= st_crs(air_shapes)) %>%
  mutate(x_center = map_dbl(geometry, ~st_point_on_surface(.x)[[1]]),
         y_center = map_dbl(geometry, ~st_point_on_surface(.x)[[2]])) %>%
  sf_to_df(fill = T)  %>%
  janitor::clean_names()

air_shapes = air_shapes %>%
  sf_to_df(fill = T) %>%
  janitor::clean_names() %>%
  mutate(geo_join_id = as.numeric(id)) %>%
  filter(id != "0")
```

# Example map

The map shows the 42 UHF neighborhoods (colors), with the zip code bounds drawn over the colors.

```{r example_map}
ggplot() + geom_polygon(data = air_shapes, 
                        aes(x = x, y = y, group = id), 
                        fill = "white", color = "black") + 
  theme_void() + geom_point(data = zip_shapes, 
                            aes(x = x_center, y = y_center, 
                                color = neighborhood)) + 
  theme(legend.position = "none")
```

# Joining Air Data with UHF Shape files 

Before joining, I think you should make `df_air` smaller (by choosing a specific year and name to track).

```{r}
#df_air_shapes = df_air %>%
#  inner_join(air_shapes, by = "id")
```

# Joining with zip code data

Joining these will be hard because the neighborhoods don't have the same names. We don't have to join them for the maps, but we'd need to join them for regression.

```{r why_joining_will_be_hard}
sort(unique(df_air$geo_place_name))

sort(unique(zip_codes$neighborhood))
```

# Exporting air and zip shape files

```{r shape_export}
write_csv(air_shapes, "cleaned_shapes/air_shapes.csv")

write_csv(zip_shapes, "cleaned_shapes/zip_shapes.csv")
```

